The \project{unRAVE} catalog

### 2016-09-08 article/unrave.tex
text edits

\subsection{Joining the models}
\label{sec:joining-the-models}

We have derived labels for all \Nspectra\ \rave\ spectra using the three models
described in previous sections.  The results from the joint model --- that includes 
the main-sequence, sub-giant and red giant branch --- shows that a single 3-label quadratic
model is too simple for the \rave\ spectral range.  The other models have problems,
too: unrealistic over-densities in label space show that the main-sequence model and
the giant model make very poor extrapolations for stars outside their respective
training sets.  For these reasons we were forced to exclude or severely penalize 
incorrect results from both models.  We emphasize that the choices here are entirely
heuristic, and depart from interpreting \thecannon\ output labels as the maxima
of individual likelihood functions.  Each model produces estimates of the labels
for a given star, and we use those estimates to produce a unified estimate, but 
this joint estimate is calculated by disregarding the probabilistic attributes
of individual estimates.

Before attempting to join the results from different models, we excluded results
in either model that had a reduced $\chi_{r}^2 > 3$.  We further discarded stars with
labels that are outside the extent of the training set.  Specifically for the
results from the giant model we (conservatively) excluded stars with derived 
$\logg > 3.5$, and for the results from the main-sequence model we excluded 
sub-giant stars ($\logg < 4$ and $\teff < 5000$~K) that were outside the 
two-dimensional ($\teff$, $\logg$) convex hull of the training set used for the main-sequence model.  Unfortunately these restrictions did not remove all spurious
results.  The reason for this can be explained with an example:  consider that our 
giant star model was trained with only giant stars but tested with both giant 
stars and dwarf stars.  Some classes of stars (e.g., metal-poor dwarfs) can 
project into a region of label space that would suggest it is a giant 
(e.g., a clump star).  These objects could have relatively low $\chi_{r}^2$ values
(e.g., $\chi_{r}^2 < 3$) and in this example, they would appear as bonafide red clump 
stars.  These incorrect projections are extrapolation errors in high dimensions that
project to `normal' parts of the label space in two dimensions.  For these reasons 
we also made use of the joint model in Section~\ref{sec:a-simple-model}
to inform whether we should adopt results from: the red giant branch model; the
main-sequence/sub-giant model; or a linear combination of the two.

In Figure \ref{fig:joint-model-differences} we show the differences in effective temperature
$\teff$ and surface gravity $\logg$ between: the main-sequence model and the joint
model; and the differences between the red giant branch model and the joint model. 
We have scaled the differences in $\teff$ and $\logg$ to make the central peak 
near $(0, 0)$ to be approximately isotropic by setting
	$\delta_{T_{\rm eff}} = 90$~K for the main-sequence model, 
	$\delta_{T_{\rm eff}} = 50$~K for the giant model, and 
	$\delta_{\log{g}} = 0.15$~dex for both models.
It is important to note that these $\delta$ values do not represent any kind of intrinsic
uncertainty or precision: they are merely normalization factors.  Adopting
substantially different scaling factors would produce in clear inconsistencies in our
results (e.g., sub-giants being misclassified as giants).  Therefore we chose these 
factors empirically to make the distributions in Figure \ref{fig:joint-model-differences}
approximately isotropic, and to some extent, comparable.  In Figure 
\ref{fig:joint-model-differences} the stars within the peak at $(0, 0)$ represent objects where
the joint model and the comparison model both report similar labels.  
The artefacts 
seen in the Hertzsprung-Russell diagrams in Figure \ref{fig:test-set-density} are
also present in Figure \ref{fig:joint-model-differences} as over-densities far away
from the central peak.  Therefore, we can 
adopt the scaled distance in labels $\teff$ and $\logg$ from the joint model to the 
main-sequence model $d_{ms}$,
\begin{eqnarray}
	d_{ms} & = & \left(\frac{T_{{\rm eff},ms} - T_{{\rm eff},joint}}{\delta_{T_{{\rm eff},ms}}}\right)^2 + \left(\frac{\log{g}_{ms} - \log{g}_{joint}}{\delta_{\log{g},ms}}\right)^2 \quad .
\end{eqnarray}

However when adopting a similar distance metric for the giant model, we found that the
giant model and the joint model produced similar label estimates (with low $\chi_r^2$
values) for metal-poor ($[{\rm Fe/H]} < -1$) dwarf stars near the turn-off.  The similarity
in estimated labels occurred in $\teff$, $\logg$, \emph{and} [Fe/H], implying that we
could not use the difference between $[{\rm Fe/H}]_{giant} - [{\rm Fe/H]}]_{joint}$
as a term that would benefit the main-sequence model.  As a result, our joint estimate 
for metal-poor dwarf stars that overlap with high-resolution studies
\citep{Reddy_2003,Reddy_2006,Bensby_2014,Valenti_Fischer_2005},
would appear as metal-poor giant stars, because the giant star model would have a much
smaller distance metric than the main-sequence model.  It is reasonable to assert that
this issue may be complicated by the increased complexity, or flexibility, of the giant
star model relative to the main-sequence model.  In any case, we found that by empirically
penalizing the giant model distance metric by absolute metallicity,
\begin{eqnarray}
	d_{giant} & = & \left(\frac{T_{{\rm eff},giant} - T_{{\rm eff},joint}}{\delta_{T_{{\rm eff},giant}}}\right)^2 + \left(\frac{\log{g}_{giant} - \log{g}_{joint}}{\delta_{\log{g},giant}}\right)^2  + \lvert[{\rm Fe/H}]_{joint}\rvert  \quad ,
\end{eqnarray}

>>>>

\subsection{Deriving joint estimates from all models}
\label{sec:joining-the-models}

We have derived labels for all \Nspectra\ \rave\ spectra using the three models
described in previous sections.  The results from our first model 
(Section~\ref{sec:a-simple-model}) --- which includes the main-sequence, sub-giant 
and red giant branch --- shows that a single 3-label quadratic model is too simple 
for the \rave\ spectral range.   The other models have problems, too: unrealistic 
over-densities in label space show that the main-sequence model and the giant model
make very poor extrapolations for stars outside their respective training sets.  
For these reasons we were forced to exclude or severely penalize incorrect results
from both models.  We emphasize that the choices here are entirely heuristic, 
and depart from interpreting \thecannon\ output labels as the maxima of individual 
likelihood functions.  Each model produces estimates of the labels for a given star, 
and we use those estimates to produce a unified estimate, but this joint estimate 
is calculated by disregarding the probabilistic attributes of individual estimates.

Before attempting to join the results from the models in Sections 
\ref{sec:unevolved-star-model} and \ref{sec:evolved-star-model}, we excluded results
in either model that had a reduced $\chi_{r}^2 > 3$.  We further discarded stars with
labels that are outside the extent of the training set.  Specifically for the
results from the giant model we (conservatively) excluded stars with derived 
$\logg > 3.5$, and for the results from the main-sequence model we excluded 
sub-giant stars ($\logg < 4$ and $\teff < 5000$~K) that were outside the two-dimensional 
($\teff$, $\logg$) convex hull of the training set used for the main-sequence model.  
Unfortunately these restrictions did not remove all spurious results.  The reason 
for this can be explained with an example:  consider that our giant star model was 
trained with only giant stars but tested with both giant stars and dwarf stars.  
Some classes of stars (e.g., metal-poor dwarfs) can project into a region of label 
space that would suggest it is a giant (e.g., a clump star).  These objects could 
have relatively low $\chi_{r}^2$ values (e.g., $\chi_{r}^2 < 3$) and in this example,
they would appear as bonafide red clump stars.  These incorrect projections are 
extrapolation errors in high dimensions that project to `normal' parts of the label 
space in two dimensions.  For these reasons we also made use of the simple model in 
Section~\ref{sec:a-simple-model} to inform whether we should adopt results from: 
the red giant branch model; the main-sequence/sub-giant model; or a linear combination 
of the two.

In Figure \ref{fig:joint-model-differences} we show the differences in effective 
temperature $\teff$ and surface gravity $\logg$ between: the main-sequence model 
(Section~\ref{sec:unevolved-star-model}) and the simplest model (Section~
\ref{sec:a-simple-model}); and the differences between the red giant branch model 
(Section~\ref{sec:evolved-star-model}) and the simplest model (Section 
\ref{sec:a-simple-model}).  We have scaled the differences in $\teff$ and $\logg$ 
to make the central peak near $(0, 0)$ to be approximately isotropic by setting
	$\delta_{T_{\rm eff}} = 90$~K for the main-sequence model, 
	$\delta_{T_{\rm eff}} = 50$~K for the giant model, and 
	$\delta_{\log{g}} = 0.15$~dex for both models.
It is important to note that these $\delta$ values do not represent any kind of 
intrinsic uncertainty or precision: they are merely normalization factors.  Adopting
substantially different scaling factors would produce in clear inconsistencies in our
results (e.g., sub-giants being misclassified as giants).  Therefore we chose these 
factors empirically to make the distributions in Figure \ref{fig:joint-model-differences}
approximately isotropic, and to some extent, comparable.  In Figure 
\ref{fig:joint-model-differences} the stars within the peak at $(0, 0)$ represent 
objects where the simple model and the comparison model both report similar labels.  
The artefacts seen in the Hertzsprung-Russell diagrams in Figure \ref{fig:test-set-density} 
are also present in Figure \ref{fig:joint-model-differences} as over-densities far away
from the central peak.  Therefore, we can adopt the scaled distance in labels 
$\teff$ and $\logg$ from the joint model to the main-sequence model $d_{ms}$,
\begin{eqnarray}
	d_{ms} & = & \left(\frac{T_{{\rm eff},ms} - T_{{\rm eff},joint}}{\delta_{T_{{\rm eff},ms}}}\right)^2 + \left(\frac{\log{g}_{ms} - \log{g}_{joint}}{\delta_{\log{g},ms}}\right)^2 \quad .
\end{eqnarray}

\stub{Warning: The metallicity correction may disappear -- I am trying to fix that}
However when adopting a similar distance metric for the giant model, we found that the
giant model and the joint model produced similar label estimates (with low $\chi_r^2$
values) for metal-poor ($[{\rm Fe/H]} < -1$) dwarf stars near the turn-off.  The similarity
in estimated labels occurred in $\teff$, $\logg$, \emph{and} [Fe/H], implying that we
could not use the difference between $[{\rm Fe/H}]_{giant} - [{\rm Fe/H]}]_{joint}$
as a term that would benefit the main-sequence model.  As a result, our joint estimate 
for metal-poor dwarf stars that overlap with high-resolution studies
\citep{Reddy_2003,Reddy_2006,Bensby_2014,Valenti_Fischer_2005},
would appear as metal-poor giant stars, because the giant star model would have a much
smaller distance metric than the main-sequence model.  It is reasonable to assert that
this issue may be complicated by the increased complexity, or flexibility, of the giant
star model relative to the main-sequence model.  In any case, we found that by empirically
penalizing the giant model distance metric by absolute metallicity,
\begin{eqnarray}
	d_{giant} & = & \left(\frac{T_{{\rm eff},giant} - T_{{\rm eff},joint}}{\delta_{T_{{\rm eff},giant}}}\right)^2 + \left(\frac{\log{g}_{giant} - \log{g}_{joint}}{\delta_{\log{g},giant}}\right)^2  + \lvert[{\rm Fe/H}]_{joint}\rvert  \quad ,
\end{eqnarray}
